A fast algorithm for computing the best response against the unplanned mixed strategies of an opponent in two-person zero-sum sequential games

### 2014-07-23 paper.tex
typos
more typos
comestics

\begin{document}
%\onecolumn
\title{Primal-dual algorithm for computing best response against the unplanned mixed strategies of an opponent in two-person zero-sum sequential games}

@@

\begin{abstract}
% LESS THAN 200 WORDS !!!!!
% 
% While medical imaging typically provides massive amounts of
% data, the extraction of relevant information in a given
% applicative context remains a difficult challenge.

% XXX : don't call it SPM but just "predictive regions or predictive
% maps"
In this manuscript, we consider the problem of computing the best response against the  unplanned mixed strategies of an opponent
in two-player zero-sum sequential games. An unplanned mixed strategy is simply a strategy chosen from a probability simplex.
The proposed technique employs the primal-dual scheme of A. Chambolle and T. Pock to solve the corresponding saddle-point problem.
The resulting algorithm is simple (all resolvent operators can be effectively computed quasi closed-form, using only elementary algebraic operations)
 and is orders of magnitude more efficient than state-of-the-art algorithms like those based on Linear Programming, Interior-Point methods,
and more recently, the Nesterov Excessive Gap Technique. The problem considered here encompasses the problem of computing the best response against an opponent's fixed strategy.
We conclude by exhibiting results on Texas Hold'em Poker.
\end{abstract}

@@

\section{Introduction: Statement of the problem}
\label{sec:intro}
%% Let $A \in \mathbb{R}^{m,n}$ be the payoff matrix; let $E \in \mathbb{R}^{r,n}$ and $e \in \mathbb{R}^r$ be a matrix so that player 1 wishes only to use those strategies  $x \in \mathbb{R}_{+}^n$ for which $Ex = e$, i.e $x \in \mathbb{R}_{+}^n$
%%  %% from player \textbf{row} perspective, so that player \textbf{row} gains $A_{i,j}$ and player \textbf{column} forfeits $A_{i,j}$ euros if the former plays the $i$th row and the former players the $j$th column.

%% We denote by $\mathbf{y} \in \mathbb{R}^{n}$ the targets to
%% be predicted, and $\mathbf{X}\in\mathbb{R}^{n \times p}$ the
%% brain images related to the presentation of different
%% stimuli. $p$ is the number of voxels and $n$ the number of
%% samples (images). Typically, $p \sim 10^3-10^5$ (for a whole volume),
%% while $n \sim 10-10^3$.  Let $\Omega \subset \mathbb{R}^3$ be
%% the 3D image domain, discretized on a finite grid. The coefficients
%% $\mathbf{w}$ define a spatial map in $\mathbb{R}^p$. Its gradient at a
%% voxel $j \in \Omega $ reads\footnote{Where it is understood that the feature vector $w$ is ``unmasked''
%% into a full 3D volume before the spatial operators are applied.}:
%% \begin{eqnarray}
%%   \mathbf{\nabla} w(j) := [\nabla_{x} w(j),
%%     \nabla_{y} w(j), \nabla_{z} w(j)] \in
%%   \mathbb{R}^3,
%% \end{eqnarray}
%% where $\nabla_{u}$ is the spatial difference operator along
%% the $u$-axis. Thus $\mathbf{\nabla}$
%% is a linear operator $\in \mathbb{R}^{3p \times
%%   p}$ with adjoint $\mathbf{\nabla} ^T = -\dive \in \mathbb{R}^{p
%%   \times 3p}$. Let $\tilde{\nabla} := [(1 - \rho)\nabla^T, \rho I]^T$ be the identity-augmented spatial gradient operator. Finally, let us recall the definition of the $(q, 1)$ mixed-norm\footnote{Where $q$ itself is allowed to be a mixed-norm.}:
%% \begin{equation}
%% \|v\|_{q, 1} := \sum_{j}{\|v_{j}\|_{q}}
%% \end{equation}

Consider a two-person zero-sum sequential game with payoff matrix $A \in \mathbb{R}^{m,n}$. Consider the problem of
finding the best response strategy $x^* \in \mathbb{R}_{+}^n \cap E^{-1}(\{e\})$ for player 1, given a fixed
(possibly mixed) strategy $y_0$ for player 2, where $E \in \{-1, 0, +1\}^{r,n}$ and $e \in \mathbb{R}^r$ encode player
1's game plan, consistent with the rules of the game. In the language of convex-optimization, one can readily give the following saddle-point formulation for this problem. Viz,

@@

where $\Delta_{m}^0$ is the singleton $\{y_0\}$ if $y_0$ is known, otherwise it is the $m-$dimensional probability simplex $\Delta_{m} := \{x \in \mathbb{R}_{+}^m \text{ s.t }\textbf{1}^Tx = 1\}$.\\\\
Noting that $i_{E^{-1}(\{e\})}(x) = \underset{\eta \in \mathbb{R}^r}{max}\text{ }\eta^T(e - Ex)$, one can re-write (1) as:
\begin{equation}
  \underset{x \in \mathbb{R}^n}{min}\text{ }\underset{y \in \mathbb{R}^m, \eta \in \mathbb{R}^r}{max}\text{ }[y,\eta]^T\left[ \begin{array}{@{}cc@{}}
    A \\ \hline
    -E
    \end{array} \right]x + G(x) - F^*(y, \eta)
\end{equation}
where:
\begin{equation}
  \begin{aligned}
    G:=i_{\mathbb{R}_{+}^n}\\
    F^*: (y, \eta) \mapsto (i_{\Delta_{m}^0}(y) - e^T\eta)
  \end{aligned}
\end{equation}
with resolvent operators given by
\begin{equation}
  \begin{aligned}
    (1 + \tau \partial G)^{-1}: x \mapsto \Pi_{\mathbb{R}_{+}^n}(x) = (x)_+\\
    (1 + \sigma \partial F^*)^{-1}: (y, \eta) \mapsto (\Pi_{\Delta_{m}^0}(y), \eta + \sigma e)% (z, \eta) \mapsto \bigl(y, \frac{\eta}{1 + \sigma}\bigr)
  \end{aligned}
\end{equation}
This way, we obtain following algorithm which solves (2).

\begin{algorithm}[htb]
\caption{Primal-dual algorithm for computing best response against unplained mixed strategies of opponent}%%  \textbf{Let}
\textbf{Initialize} $\tilde{x}^{(0)} = x^{(0)} \in \mathbb{R}^n$; $y^{(0)} \in \mathbb{R}^{m}$; $\eta^{(0)} \in \mathbb{R}^{m}$;
$\tau, \sigma > 0 \text{ s.t. }\tau\sigma (\|A\|^2 + \|E\|^2) \le 1$; $k = 0$.\\
\Repeat{
  convergence}{
  \begin{eqnarray*}
    y^{(k+1)} &\leftarrow& \Pi_{\Delta_{m}^0}\bigl(y^{(k)} + \sigma A \tilde{x}^{(k)}\bigr)\\
    \eta^{(k+1)} &\leftarrow& \eta^{(k)} - \sigma E \tilde{x}^{(k)} + \sigma e\\
    x^{(k+1)} &\leftarrow& (x^{(k)} - \tau A^Ty^{(k+1)} + E^T\eta^{(k+1)})_+\\
    \tilde{x}^{(k+1)} &\leftarrow& 2x^{(k+1)} - x^{(k)}\\
    k &\leftarrow& k + 1
  \end{eqnarray*}
} \Return $x^{(k)}$
\label{Tab:pseudocode_lbfgs}
\end{algorithm}

>>>>

\begin{document}
%\onecolumn
\title{A fast algorithm for computing the best response against the unplanned mixed strategies of an opponent in two-person zero-sum sequential games}

@@

\begin{abstract}
% LESS THAN 200 WORDS !!!!!
% 
% While medical imaging typically provides massive amounts of
% data, the extraction of relevant information in a given
% applicative context remains a difficult challenge.
% XXX : don't call it SPM but just "predictive regions or predictive
% maps"
In this manuscript, we consider the problem of computing the best response against the  unplanned mixed strategies of an opponent
in two-player zero-sum sequential games. The problem considered here encompasses the problem of computing the best response against an opponent's fixed strategy.
The problem considered here encompasses the problem of computing the best response against an opponent's fixed strategy.
An unplanned mixed strategy is simply a strategy chosen from a probability simplex.
The proposed technique employs the primal-dual scheme of A. Chambolle and T. Pock to solve the corresponding saddle-point problem.
The resulting algorithm is simple: all resolvent operators can be effectively computed quasi closed-form, using only elementary algebraic operations.
%%  and is orders of magnitude more efficient than state-of-the-art algorithms like those based on Linear Programming, Interior-Point methods,
%% and more recently, the Nesterov Excessive Gap Technique.
By way of example, we conclude by exhibiting results on Texas Hold'em Poker.
\end{abstract}

@@

\section{Introduction: Statement of the problem}
\label{sec:intro}

To begin, let us introduce some technical vocabulary we will be using in this paper. The reader should lookup any standard textbook
on convex optimization for a tutorial introduction to these notions. Viz,\\
$\mathbb{R}_+$: \quad \textit{the nonnegative part of the real-line $\mathbb{R}$}\\
$(x)_+$: \quad \textit{the component-wise maximum of the vector $x$ and 0}\\
$\Delta_{m}$: \quad \textit{the $m-$dimensional probability simplex
%$\{x \in \mathbb{R}_{+}^m\text{ }| \text{ }\textbf{1}^Tx = 1\}$
}\\
$\Pi_C$: \quad \textit{the euclidean projector onto the convex set $C$}\\
$\|K\|$: \quad \textit{the operator norm of a matrix $K$}\\
$F^*$: \quad \textit{the Fenchel transform of a convex l.s.c\footnote{lower semi-contnuous} function $F$}\\
$E^{-1}(U)$: \quad \textit{the pre-image of $U \subset \mathbb{R}^n$ under a matrix $E \in \mathbb{R}^{r,n}$, namely the set $\{x \in \mathbb{R}^n\text{ }| \text{ }Ex \in U\}$}\\\\
Now, consider a two-person zero-sum sequential game with payoff matrix $A \in \mathbb{R}^{m,n}$. We are interested in the problem of
finding the best response strategy $x^* \in \mathbb{R}_{+}^n \cap E^{-1}(\{e\})$ for player 1, given a fixed
(possibly mixed) strategy $y_0$ for player 2, where $E \in \{-1, 0, +1\}^{r,n}$ and $e \in \mathbb{R}^r$ encode player
1's game plan, consistent with the rules of the game. In the language of convex-optimization, one can readily give the following saddle-point formulation for this problem. Viz,

@@

where $\Delta_{m}^0$ is $\Delta_{m}$ if $y_0$ is known, otherwise it is the singleton
%% \footnote{Note that the origin $\textbf{0} \in \mathbb{R}^m$ can be seen a degenerate simplex with radius $0^+$,
%% whilst a singleton  $\{y_0\}$ can be seen as a $y_0-$translation of this simplex.}
$\{y_0\}$.\\\\

In section II, we give a brief overview of existing methods for solving (1). We elaborate our proposed primal-dual algorithm in section III.

\section{Related work}
pending...

\section{The algorithm}
Noting that $i_{E^{-1}(\{e\})}(x) = \underset{\theta \in \mathbb{R}^r}{max}\text{ }\theta^T(e - Ex)$, one can re-write (1) as:
\begin{equation}
  \underset{x \in \mathbb{R}^n}{min}\text{ }\underset{y \in \mathbb{R}^m, \theta \in \mathbb{R}^r}{max}\text{ }
  \left[\begin{array}{@{}cc@{}}y\\\theta\end{array}\right]^TKx + G(x) - F^*(y, \theta)
\end{equation}
where:
\begin{equation}
  \left.
  \begin{aligned}
    K := \left[\begin{array}{@{}cc@{}}A\\-E\end{array}\right]\\
    G:=i_{\mathbb{R}_{+}^n}\\
    F^*: (y, \theta) \mapsto (i_{\Delta_{m}^0}(y) - e^T\theta)
  \end{aligned}
  \right\}
\end{equation}

Furthermore, $F^*$ and $G$ have resolvent (a.k.a \textit{proximity}) operators given by the simple formulae:
\begin{equation}
  \left.
  \begin{aligned}
    (1 + \tau \partial G)^{-1}: x \mapsto \Pi_{\mathbb{R}_{+}^n}(x) = (x)_+\\
    (1 + \sigma \partial F^*)^{-1}: (y, \theta) \mapsto (\Pi_{\Delta_{m}^0}(y), \theta + \sigma e)% (z, \theta) \mapsto \bigl(y, \frac{\theta}{1 + \sigma}\bigr)
  \end{aligned}
  \right\}
\end{equation}
This way, we obtain following primal-dual algorithm for solving (2).

\begin{algorithm}[htb]
\caption{Primal-dual algorithm for computing best response against unplained mixed strategies of opponent}%%  \textbf{Let}
\textbf{Initialize} $\tilde{x}^{(0)} = x^{(0)} \in \mathbb{R}^n$; $y^{(0)} \in \mathbb{R}^{m}$; $\theta^{(0)} \in \mathbb{R}^{r}$;
$\tau, \sigma > 0 \text{ s.t. }\tau\sigma \|K\|^2 \le 1$; $k = 0$.\\
\Repeat{
  convergence}{
  \begin{eqnarray*}
    y^{(k+1)} &\leftarrow& \Pi_{\Delta_{m}^0}\bigl(y^{(k)} + \sigma A \tilde{x}^{(k)}\bigr)\\
    \theta^{(k+1)} &\leftarrow& \theta^{(k)} - \sigma E \tilde{x}^{(k)} + \sigma e\\
    x^{(k+1)} &\leftarrow& \bigl(x^{(k)} - \tau A^Ty^{(k+1)} + E^T\theta^{(k+1)}\bigr)_+\\
    \tilde{x}^{(k+1)} &\leftarrow& 2x^{(k+1)} - x^{(k)}\\
    k &\leftarrow& k + 1
  \end{eqnarray*}
} \Return $x^{(k)}$
\label{Tab:pseudocode_lbfgs}
\end{algorithm}
